---
title: "Peer-graded Assignment: Course Project 1"
output: html_document
---

```{r setup, include=FALSE}
library("knitr")

knitr::opts_chunk$set(echo = T, warning = F, error = T)
knit2html("")
```

## 1. Code for reading in the dataset and/or processing the data

```{r}
 activity <- read.csv("C:/Users/auke.beeksma/Documents/activity.csv")

```

## 2. Histogram of the total number of steps taken each day

```{r}
library(dplyr)
activity  <- activity[complete.cases(activity), ]

activityday <- tapply(activity$steps, activity$date, sum, na.rm = T)
hist(activityday, main = "Steps a day")
```


## 3. Mean and median number of steps taken each day

```{r}
steps <- activity %>% 
  group_by(date) %>% 
  summarize_all(sum)

mean(steps$steps)
median(steps$steps)

```

## 4. Time series plot of the average number of steps taken

```{r}


plot(steps$date, steps$steps, xlab = "Day in october/november", ylab = "Average step a day")
abline( h = mean(steps$steps), col = "blue")
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps

```{r}

stepsmax5 <- activity %>% 
  group_by(interval) %>% 
  summarize_all(mean)
 

which.max(stepsmax5$steps)
stepsmax5[104,]
```

## 6. Code to describe and show a strategy for imputing missing data
Strategy: replace missing data with imputed "mean" data. This generates a random sample of numbers between zero and 16492. This forms a better representation of the data then for example "mean" because of the many values of zero in the data. The impact of adding ''random numbers'' seems little (only a small difference in the mean)

```{r}

library(Hmisc)
activity <- read.csv("C:/Users/auke.beeksma/Documents/activity.csv")
nrow(activity) - sum(complete.cases(activity))

activity$imputed_steps <- with(activity, impute(steps, "random"))

steps <- activity %>% 
  group_by(date) %>% 
  summarize_all(sum)

mean(steps$imputed_steps)
median(steps$imputed_steps)


```


## 7. Histogram of the total number of steps taken each day after missing values are imputed

```{r}

activity <- tapply(activity$imputed_steps, activity$date, sum, na.rm = T)
hist(activity, main = "Steps a day")
```

## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}

library(lubridate)
activity <- read.csv("C:/Users/auke.beeksma/Documents/activity.csv")

activity <- activity %>% 
  select(steps, date, interval) %>% 
  mutate(weekday = wday(date, label = T)) 

weekday <- filter(activity, !weekday %in% c("za","zo"))
weekend <- filter(activity, weekday %in% c("za","zo"))

weekday  <- weekday[complete.cases(weekday), ]
weekend  <- weekend[complete.cases(weekend), ]

par(mfrow=c(2,1), col = "blue")


weekday %>%
  group_by(interval) %>% 
  summarise (mean(steps)) -> meanstepw
  
  plot(meanstepw, type = "l", ylim=range(c(0,250)), main = "Steps per 5 minute interval on daily basis week, 0:00-23:59")

weekend %>%
  group_by(interval) %>% 
  summarise (mean(steps)) -> meanstepwk

plot(meanstepwk, type = "l", ylim=range(c(0,250)), main = "Steps per 5 minute interval on daily basis weekend, 0:00-23:59")

```


